---
title: "Regression Models Course Project"
author: "Adam Ficke"
date: "5/9/2020"
output: html_document
---

## Executive Summary 

This paper is interested in addressing the relationship between car transmission type and fuel economy. Specitically, it answers: 


* “Is an automatic or manual transmission better for MPG”
* "Quantify the MPG difference between automatic and manual transmissions"


The paper contains 5 sections: 

1. Data Cleaning 
2. Data exploration 
3. Variable Reduction
4. Significance Tests 
5. Results 

***
### Data Cleaning 

```{r include=FALSE}
require(dplyr)
require(tidyverse)
require(ggplot2)
require(corrplot)
require(RColorBrewer)
require(PerformanceAnalytics)
require(GGally)
require(caret)
require(AppliedPredictiveModeling)
require(glmnet)
require(bestglm)
require(doParallel)

data("mtcars")
mtcars.fac <-
  mtcars %>% mutate_at(c("cyl", "vs", "am", "gear", "carb"), as.factor)

```


First we'll read in our data and adjust the data types as needed
```{r, echo = FALSE}

as.matrix(sapply(mtcars, class))
as.matrix(sapply(mtcars.fac, class))

```

### Data Exploration

First we'll examine the response data:

```{r, examine response data 1, echo = FALSE}
qplot(mpg, data = mtcars, binwidth = 1,main = "Distribution of MPG Data") + 
    theme(plot.title = element_text(hjust = 0.5))
```

Next we'll look at the superficial relationship between transmission type and fuel economy:
```{r, examine response data 2, echo = FALSE}
P1 <- ggplot(data = mtcars.fac,
             aes(x = am, y = mpg)) +
   geom_boxplot() + 
  ggtitle("Transmission Type vs MPG (0 = automatic, 1 = manual)") + 
  theme(plot.title = element_text(hjust = 0.5))
P1
```

Now we'll examine the corelations in our data. Since there are few predictors, we'll look at all of them: 
```{r, examine response data, echo = FALSE}
Cars.Cor <- cor(mtcars)
corrplot(
  Cars.Cor,
  type = "upper",
  order = "hclust",
  col = brewer.pal(n = 5, name = "RdYlBu")
)
```


###  Variable Reduction 

Here we'll go through a series of models to 
see if the superficial relationship we found 
in the first step holds







